---
title: "Class 14"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(BiocManager)
library(DESeq2)
```

```{r}
counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
head(counts)
head(metadata)

# Number of genes in total
nrow(counts)
```
```{r}
View(metadata)
```

```{r}
# Check whether names are right 
colnames(counts) == metadata$id
```

## Section 3. Walk Through of gene expression analysis 
```{r}
control.inds <- metadata$dex == "control"
control <- metadata[control.inds, ]
control <- metadata[metadata[,"dex"]=="control",]
head(counts[, control$id])
control.mean <- rowSums( counts[ ,control$id] )/ length(control$id)
names(control.mean) <- counts$ensgene

treated.inds <- metadata$dex == "treated"
treated <- metadata[treated.inds, ]
treated <- metadata[metadata[,"dex"]=="treated",]
head(counts[, treated$id])
treated.mean <- rowSums( counts[ ,treated$id] )/ length(treated$id)
```

### Combine two data table
```{r}
meancounts <- data.frame(control.mean, treated.mean)
colSums(meancounts)
plot(meancounts$control.mean, meancounts$treated.mean, xlab = "Control", ylab = "Treated")
plot(meancounts$control.mean, meancounts$treated.mean, 
     xlab = "Control", ylab = "Treated", log = "xy")
```

### Data Transformation
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)

# Remove zero
to.rm <- unique(which(meancounts[, 1:2] == 0, arr.ind = T)[,1])
```

```{r}
mycounts <- meancounts[-to.rm,]
nrow(mycounts)
```

### How many genes are up in the drug treated cells
```{r}
up.inds <- mycounts$log2fc > 2
sum(up.inds)

down.inds <- mycounts$log2fc < -2
sum(down.inds)
```

```{r}
head(mycounts[up.inds,])
```


## 4. Adding annotation data
```{r}
anno <- read.csv("data/annotables_grch38.csv")
head(anno)
```
```{r}
# Use the merge function to add the annotation data
mycounts.anno <- merge(mycounts, anno, by.x = "row.names", by.y = "ensgene")
mycounts.anno 
```

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```

```{r}
columns(org.Hs.eg.db)
```

### We will use the **mapIds()** function from bioconductor to add annotation data to our results
```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db, keys = row.names(mycounts), 
                          keytype = "ENSEMBL",
                          column = "SYMBOL")    

mycounts$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")
```

```{r}
head(mycounts)
```

## 5.DESeq2 analysis
```{r}
library(DESeq2)
citation("DESeq2")
```
```{r}
dds <- DESeqDataSetFromMatrix(countData= counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```
```{r}
dds <- DESeq(dds)
res <- results(dds)
head(res)
```
```{r}
summary(res)
```

```{r}
res05 <- results(dds, alpha = 0.05)
summary(res05)
```

## Save result as a data frame
```{r}
res01 <- as.data.frame(results(dds, alpha = 0.01))
head(res01)
```

## Sort results by the adjusted p-value
```{r}
x <- c(5,4,1,2)
x[order(x)]

ord.inds <- order(res01$padj)
write.csv(res01[ord.inds,], file = "signif01_results.csv")
```

## 6. Data Visualization (Volcano Plot)
```{r}
mycols <- rep("gray", nrow(res01))
mycols[res01$padj < 0.01] <- "black"
#mycols[abs(res01$log2FoldChange) > 2] <- "blue"
mycols[(res01$padj < 0.01) & (abs(res01$log2FoldChange) > 2)] <- "red"

plot(res01$log2FoldChange, -log(res01$padj), xlab = "-log Fold Change (Drug/Control)",
     ylab = "-log p-value", col = mycols)
```

