---
title: "Class11"
author: "Tianhao Qiu"
date: "5/7/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB database

The [PDB](http://www.rcsb.org/ ) is the main repository for biomolecular structure data.

Contents of PDB:

>Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and determine the percentage of structures solved by X-Ray and Electron Microscopy. From the website what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers.

```{r}
db <-read.csv('Data Export Summary.csv')
head(db)
```
```{r}
# Proportion data
db$Total / sum(db$Total) * 100

# Proportion (protein)
sum(db$Proteins)/ sum(db$Total) * 100
```

>Q2: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

```{r}
1157
# see http://www.rcsb.org/pdb/results/results.do?tabtoshow=Current&qrid=2117E9DB
```
```{r}
library(datapasta)

temp <- data.frame(stringsAsFactors=FALSE,
           Experimental.Method = c("Electron Microscopy", "Multi Method", "NMR",
                                   "Other", "X-Ray", "Total"),
                      Proteins = c(2277, 129, 11062, 256, 126880, 140604),
                 Nucleic.Acids = c(31, 5, 1279, 4, 2012, 3331),
            ProteinComplex = c(800, 2, 259, 6, 6547, 7614),
                         Other = c(0, 1, 8, 13, 8, 30),
                         Total = c(3108, 137, 12608, 279, 135447, 151579)
        )

```

```{r}
library(bio3d)
example(plot.bio3d)
```

```{r}
pdb <- read.pdb("1hsg")
pdb
attributes(pdb)
```

> Q6. How many amino acid residues are there in this pdb object and what are the two non- protein residues?

``` ={r}
198, HOH, MK1
```

>Q7. What type of R object is ?

```{r}
str(pdb$atom)
typeof(pdb$atom)
```


```{r}
head(pdb$atom)

# Print a subset of $atom data for the first two atoms  
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]


# Note that individual $atom records can also be accessed like this  
pdb$atom$elety[1:2]

# Which allows us to do the following  
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab= "B-factor")
```

## Atom selection (atom.select())

```{r}
prot.pdb <- atom.select(pdb, "protein", value = T)
ligd.pdb <- atom.select(pdb, "ligand", value = T)
write.pdb(ligd.pdb, file ="1hsg_ligand.pdb")
write.pdb(prot.pdb, file ="1hsg_protein.pdb")
```

```{r}
atom.select(pdb, "ligand", value = T)
```

```{r}

# Select all C-alpha atoms (return their indices)  
ca.inds <- atom.select(pdb, "calpha") 
ca.inds

# Print details of the first few selected atoms  
head( pdb$atom[ca.inds$atom, ] )

# And selected xyz coordinates  
head( pdb$xyz[, ca.inds$xyz] )

# Select chain A 
a.inds <- atom.select(pdb, chain="A") 
 
# Select C-alphas of chain A 
ca.inds <- atom.select(pdb, "calpha", chain="A") 
 
# We can combine multiple selection criteria to return their intersection 
cab.inds <- atom.select(pdb, elety=c("CA","CB"), chain="A", resno=10:20)
```

```{r}
# Download some example PDB files 
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A") 
files <- get.pdb(ids, split = TRUE)
```

```{r}
# Extract and align the chains we are interested in  
pdbs <- pdbaln(files, fit = TRUE)
 
# Print to screen a summary of the 'pdbs' object  
pdbs
```

